---
title: "Class17_Vaccination"
author: "Hayoung A15531571"
date: "11/23/2021"
output:
  pdf_document:
    latex_engine: xelatex
  html_document: default
  
---

## Getting Started

Let's start by importing our data!
```{r}
# Import vaccination data
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```
```{r}
library(dplyr)
```

>> Q1. What column details the total number of people fully vaccinated?

The column "persons_fully_vaccinated" details the total number of fully vaccinated people. 

>> Q2. What column details the Zip code tabulation area? 

"zip_code_tabulation_area" column shows the zip code

>> Q3. What is the earliest date in this dataset?

```{r}
head(n=1, vax$as_of_date)
```
```{r}
dplyr::first(vax$as_of_date)
```
The earliest date is 2021-01-5

>> Q4. What is the latest date in this dataset?

```{r}
dplyr::last(vax$as_of_date)
```
The latest date is 2021-11-16


Calling skim!
```{r}
skimr::skim(vax)
```

>> Q5. How many numeric columns are in this dataset?

9

>> Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column?

There are 8256 NA values in this column

>> Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)?

```{r}
(sum( is.na(vax$persons_fully_vaccinated) ) /
NROW(vax$persons_fully_vaccinated)) *100
```
10.17% of the data is missing here

# Working with dates

```{r}
#Let's use lubridate to make our lives easier when dealing with dates and times
#install.packages("lubridate")
library(lubridate)
```

```{r}
#Specify our format
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
#Example of what we can now do easily
today() - vax$as_of_date[1]
```

>> Q9. How many days have passed between the first and last day of the report?

```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```

It has been 315 days since the first update and to the latest update

>> Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?

```{r}
NROW(unique(vax$as_of_date))
```
There are 46 unique dates in the dataset

## Working with ZIP codes

```{r}
#Load first
library(zipcodeR)
```
```{r}
geocode_zip('92037')
```

An example
```{r}
#Calculate the distance between the centroids of any two ZIP codes in miles, e.g.
zip_distance('92037','92109')
```
The distance between these two zip codes is 2.33 miles

We can pull census data about ZIP code areas (including median household income etc.). For example:
```{r}
reverse_zipcode(c('92037', "92109", "92130") )
```

```{r}
# Pull data for all ZIP codes in the dataset
zipdata <- reverse_zipcode( vax$zip_code_tabulation_area )
```


## Focus on the San Diego Area

Let’s now focus in on the San Diego County area by restricting ourselves first to vax$county == "San Diego" entries. We can use base R or dplyr

```{r}
library(dplyr)

sd <- filter(vax, county == "San Diego")

nrow(sd)
```

Using dplyr is often more convenient when we are subsetting across multiple criteria, for example all San Diego county areas with a population of over 10,000.

```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```

>> Q11. How many distinct zip codes are listed for San Diego County

```{r}
length(sd$zip_code_tabulation_area)
length(unique(sd$zip_code_tabulation_area))
```
There are 107 unique zip codes listed for the county

>> Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?

```{r}
ind <- which.max(sd$age12_plus_population)
sd[ind,]
```
The zip code is 92154, with 76365.2 people over the age of 12

What is the population of 92037 ZIP code area?
```{r}
filter(sd, zip_code_tabulation_area == "92037")[1,]
```


```{r}
sd_q13 <- filter(vax, county == "San Diego" & as_of_date == "2021-11-09")

nrow(sd_q13)
```

>> Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2021-11-09”?

```{r}
sd.now <- filter(sd, as_of_date=="2021-11-09")

mean(sd.now$percent_of_population_fully_vaccinated, na.rm = TRUE)
```
The average is 67.28%

```{r}
summary(sd.now$percent_of_population_fully_vaccinated)
```

>> Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2021-11-09”?

```{r}
hist(sd.now$percent_of_population_fully_vaccinated)
```

```{r}
library(ggplot2)

ggplot(sd.now) +
  aes(percent_of_population_fully_vaccinated) +
  geom_histogram(bins=15)
```

```{r}
filter(sd.now, zip_code_tabulation_area == "92037")
```


# Focus on UCSD / La Jolla

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")

#Age 5+ population
ucsd[1,]$age5_plus_population
```

>> Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
ggplot(ucsd) +
  aes(x=as_of_date, y=percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated")
```

>> Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2021-11-16")

head(vax.36)
```

```{r}
mean(vax.36$percent_of_population_fully_vaccinated)
```

```{r}
ggplot(ucsd) +
  aes(x=as_of_date, y=percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated") +
  geom_hline(yintercept = .6629812)
```

>> Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”?

```{r}
summary(vax.36$percent_of_population_fully_vaccinated)
```
The average is 66.30%

>> Q18. Using ggplot generate a histogram of this data.

```{r}
ggplot(ucsd) +
  aes(percent_of_population_fully_vaccinated) +
  geom_histogram()
```

>> Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
```
```{r}
vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)
```

The average overal was 66.30%, so the 92040 average is lower at 52.04%
The 92109 averave is higher at 68.77%

>> Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)


ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  labs(x="Date", y="Percent Vaccinated",
       title="Vaccination Rate Across California",
       subtitle="Only areas with a population above 36k are shown") +
  geom_hline(yintercept = 0.66, linetype="dashed")
```


