---
title: "Class_19"
author: "Hayoung A15531571"
date: "11/30/2021"
output:
  pdf_document: default
  html_document: default
---

Section 4 Homework

Let's load the file into our worksheet
```{r}
expr <- read.csv("updated.csv")
```

```{r}
head(expr)
```

>> Q13. Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes. 

```{r}
table(expr$geno)
```
There are 108 A|A, 233 A|G, and 121 G|G genotype samples.

```{r}
library(plyr)

p_meds <- ddply(expr, .(geno), summarise, med = median(exp) )
p_meds
```
The median expressions of the genotypes are 31.25 for A|A, 25.06 for A|G, and 20.07 for G|G

>> Q14. Generate a boxplot with a box per genotype, what can you infer from the relative expression value between A|A and G|G displayed in this plot? Does the SNP effect the
expression of ORMDL3?

```{r}
library(ggplot2)

plot <- ggplot(expr) + 
  aes(x = geno, y = exp, col = geno) +
  geom_boxplot()

plot
```
From the above plot, we can see that for A|A and G|G, the "boxes" do not overlap, and thus we can infer that they are relatively separate from one another, and there is a difference in expression. 
The SNP does effect the expression of ORMDL3, we can see a difference in expression when comparing genotypes. 

